---
title: "Introduction to Statistics Project"
author: "Shishir Shirbhate"
date: "July 27, 2016"
output: html_document
---

```{r}

library(readxl)
library(readr)
library("data.table")

Loan <- read.csv(file="loan_10000_new.csv",head=TRUE,sep=",")

# Empty columns contain NAs only or ""s only, they have has no variability. This code removes all columns without variability (which is probably a plus in this case).
Loan <- Loan[!sapply(Loan, function(var) (length(unique(var)) == 1))]

Loan <- na.omit(Loan)

# Get first 2000 rows
# Loan <- Loan[c(1:2000),]

# Remove 'xx' from zip code end
Loan$zip_new <- gsub("xx","",Loan$zip_code)
Loan$int_rate <- Loan$int_rate * 100
```

# Model 1

```{r}

model <- lm(formula= int_rate ~ loan_amnt + revol_util + dti + tot_cur_bal + tot_hi_cred_lim + home_ownership + total_bc_limit + term + percent_bc_gt_75 + purpose, data = Loan)

summary(model)

```

Testing the assumptions

home_ownership, term and purpose are categorical string variables
Hence removed those in checks for collinearity

```{r}

pairs(~int_rate + loan_amnt + revol_util + dti + tot_cur_bal + tot_hi_cred_lim + home_ownership + total_bc_limit + installment + term + percent_bc_gt_75 + purpose, data = Loan)

cor(Loan[,c("annual_inc","loan_amnt","revol_util","dti","tot_cur_bal","tot_hi_cred_lim","total_bc_limit","installment")])

```

'tot_cur_bal' is highly collinear with 'tot_hi_cred_lim'

Hence removing 'tot_cur_bal' (Since it has higher p-value among both).

# Model 2

After changing home_ownership to factors the other variables become insignificant. 
Removing home_ownership.

```{r}

Loan$purposeF = factor(Loan$purpose)
Loan = within(Loan, purposeF<-relevel(purposeF,ref="credit_card"))

Loan$homeF <- factor(Loan$home_ownership)
Loan <- within(Loan,homeF <- relevel(homeF, ref="NONE"))

model <- lm(formula= int_rate ~  loan_amnt + revol_util + dti + tot_hi_cred_lim +  total_bc_limit + term + percent_bc_gt_75 + purposeF, data = Loan)

summary(model)

```

# Model 3

Removing purpose as many of the variables are insignificant.
loan_amnt becomes insignificant in the above model. 
Since the scale is too high. Taking log of loan_amnt and using it in the model.

```{r}

model <- lm(formula= int_rate ~  loan_amnt + revol_util + dti + tot_hi_cred_lim +  total_bc_limit + term + percent_bc_gt_75, data = Loan)

summary(model)

```

# Final Model

```{r}
# Convert interest rate into percentages

Loan$purposeF = factor(Loan$purpose)
Loan = within(Loan, purposeF<-relevel(purposeF,ref="credit_card"))

model <- lm(formula=int_rate ~  loan_amnt + revol_util + dti + tot_hi_cred_lim +  total_bc_limit + term + percent_bc_gt_75 + purposeF, data = Loan)

summary(model)

confint(model, level = .90)
```

Check for Standardized residuals

```{r}

model.stres <- rstandard(model)
plot(model$fitted.values, model.stres, pch = 16, main = "Standardized Residual Plot", xlab = "Fitted Interest Rate", ylab = "Standardized Residuals")
abline(0,0, lty=2, col="red")

```

Check for normality 

```{r}

h <- hist(model.stres)
# x <- model.stres
# xfit <- seq(min(x), max(x), length = 50)
# yfit <- dnorm(xfit, mean = mean(x), sd = sd(x))
# yfit <- yfit*diff(h$mids[1:2])*length(x)
# lines(xfit, yfit, col="blue")

qqnorm(model.stres)
qqline(model.stres,col="red")

# shapiro.test(model.stres)

```

Plot regression results of non-factor predictos

```{r}
library(coefplot)
coefplot(model, predictors=c("loan_amnt","tot_hi_cred_lim","total_bc_limit"), xlab = "value", ylab="", title="",sort = c("magnitude")) + theme(text = element_text(size=20),axis.text.x = element_text(angle=0, vjust=1) + xlab="% change in interest rate per dollar change")

coefplot(model, predictors=c("revol_util","dti","percent_bc_gt_75"), xlab = "value", ylab="", title="",sort = c("magnitude")) + theme(text = element_text(size=20),axis.text.x = element_text(angle=0, vjust=1))

```

Plot regression results of factor predictor term - Purpose

```{r}
library(coefplot)
coefplot(model, predictors=c("purposeF"), xlab = "value", ylab="", title="",sort = c("magnitude")) + theme(text = element_text(size=20),axis.text.x = element_text(angle=0, vjust=1)) + theme (axis.title.y  = element_text(size=16))
```

Plot regression results of factor predictor term - Term

```{r}
coefplot(model, predictors=c("term"), xlab = "value", ylab="", title="", sort = c("magnitude")) + theme(text = element_text(size=20),axis.text.x = element_text(angle=0, vjust=1))
```

Descriptive Stats

```{r}

library(ggplot2)


ggplot(Loan, aes(x=loan_amnt)) + 
         geom_bar(color="blue",fill="blue") + 
         theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
    xlab("Loan Amount")


ggplot(Loan, aes(x=tot_hi_cred_lim)) + 
         geom_bar(color="blue",fill="blue") + 
         theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
    xlab("Total High credit limit")


ggplot(Loan, aes(x=total_bc_limit)) + 
         geom_bar(color="blue",fill="blue") + 
         theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
    xlab("Total Bank Card credit limit")


ggplot(Loan, aes(x=dti)) + 
         geom_bar(color="blue",fill="blue") + 
         theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
    xlab("Monthly Debt to Income Ratio (%)")

ggplot(Loan, aes(x=revol_util)) + 
         geom_bar(color="blue",fill="blue") + 
         theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
    xlab("Credit card balance / Credit limit (%)")

ggplot(Loan, aes(x=percent_bc_gt_75)) + 
         geom_bar(color="blue",fill="blue") + 
         theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
    xlab("# accounts used > 75% of limit / Total Accounts")

# To change plot order of bars, change levels in underlying factor
reorder_size <- function(x) {
  factor(x, levels = names(sort(table(x))))
}

ggplot(Loan, aes(reorder_size(home_ownership))) +
  geom_bar(color="blue",fill="blue") + 
  theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) + 
  xlab("Home Ownership")

ggplot(Loan, aes(reorder_size(purpose))) +
  geom_bar(color="blue",fill="blue") + 
  theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=0, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) +
  xlab("Purpose") + 
  coord_flip()

# Bar graph of counts
ggplot(data=Loan, aes(x=emp_length)) + 
  geom_bar(color="blue",fill="blue") + 
  theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=45, vjust=0.5, size=16)) + 
  theme(axis.title.y = element_text(face="bold", colour="#990000", size=20),axis.text.y  = element_text(angle=0, vjust=0.5, size=16)) +
  scale_x_discrete(limits=c("< 1 year","1 year","2 years","3 years","4 years","5 years","6 years","7 years","8 years","9 years","10+ years","n/a"))+
  xlab("Employment Length")



```

